---
title: "Joint factorization Yuan + GWAS"
output: html_notebook
---
Here, I would like to do a quick analysis of joint PCA on the eQTL loading matrix and the GWAS SNP z-scores.
The regression seemed to yield some hits, but I am not sure if the PCA would.
Let's take a look:

```{r setup, echo = FALSE}
pacman::p_load(data.table, tidyr, dplyr, readr, ggplot2, stringr,stats, cowplot, glmnet)
#source("./function_dossier.R")
```

#10/11

```{r}
gwas_out <-  fread("/work-zfs/abattle4/ashton/snp_networks/scratch/eqtl_gwas_reverse/gwas_extracts/snpsmp.z.tsv") %>% drop_na()
l_matrix <- fread("/work-zfs/abattle4/ashton/snp_networks/scratch/eqtl_gwas_reverse/snspMF_matrix.clean.txt") %>% rename("hg38" = V2) %>% arrange(hg38)
bed.conversion <- fread("/work-zfs/abattle4/ashton/snp_networks/scratch/eqtl_gwas_reverse/loadings.hg19.bed") %>% mutate("ids" = paste0(gsub(pattern = "chr", replacement = "", x = V1),":", V3)) %>% rename("hg38" = V4)
colnames(bed.conversion) <- c("chr", "start", "end", "hg38", "ids")

gwas_out <- left_join(gwas_out, bed.conversion, by = "ids") %>% arrange(hg38)

#we need to expand GWAS out to include all the copies
#Need to drop all the snps we don't have, maybe dropped due to something or other....
l_matrix$hg38 <- gsub(x = l_matrix$hg38, pattern = "_", replacement = ":")
filt.l <- l_matrix %>% filter((hg38 %in% gwas_out$hg38))
dim(filt.l)
```
Take a look
```{r}
any(filt.l$hg38 != gwas_out$hg38) #perfectly ordered
joint.mf.gwas <-cbind(filt.l, gwas_out %>% select(-hg38, -chr, -start, -end, -ids)) %>% select(-V1, -hg38)
joint.mf.gwas.meta <- cbind(gwas_out %>% select(hg38, chr, start, end, ids), filt.l$V1)  

#fix names
gwas_names <-fread("../fourth_pass_universal/gwas_extracts/trait.names.txt", header = FALSE)
head(gwas_names)
any(colnames(joint.mf.gwas)[24:78] != gwas_names$V1) #names lined up?
colnames(joint.mf.gwas) <- c(paste0("F", 1:23), gwas_names$V2)
head(joint.mf.gwas)
```


Okay, so now we want to run the PCA analysis.
```{r}
#note: tons (500,000 ish) of NAs in the data. Where are they?
#modification to remove at the beginning
apply(joint.mf.gwas, 2, function(x) sum(is.na(x)))
nas <- which(is.na(joint.mf.gwas$Lymphocyte_count))
head(joint.mf.gwas[nas,])
head(joint.mf.gwas.meta[nas,])
#checking these out in the original files
#Cursory look at this indicates they aren't actually present in the data. I should have dropped them from the get go you nimcompoop.

```
Psych, start with the correlation heatmap
```{r}
heatmap.joint <- sepCorHeatmap(abs(joint.mf.gwas), colnames(joint.mf.gwas), ngwas = 55, neqtl = 23, dtCluster = TRUE)
ggsave(filename = "../snspMF_gwas_analysis/abs_heatmap.png",plot = heatmap.joint)
```
NOW, we can try the PCA
```{r}
first <- fullSpectralDecomp(joint.mf.gwas,byCol = TRUE, K = 20)
first$scree
save(first, file="../snspMF_gwas_analysis/joint.pca.RData")
first$svd$v
pca.heatmap.dense <- pcaPlotMaker(first$svd$v,joint.mf.gwas,  colnames(joint.mf.gwas), "Singular Vectors, joint PCA (std)", ngwas = 55, neqtl = 23)
pca.heatmap.dense
```
Hmm, interesting. Some real overlap later on it seems.
Let's look at the contributions?
```{r}
mix.snspmf <- mixtureAnalysis(first$svd$v, colnames(joint.mf.gwas), ngwas = 55, neqtl = 23)

mix.snspmf$plot
```
Try the sparse one too
```{r}
snspmf.sparse <- sparseSpectralDecomp(joint.mf.gwas,byCol = TRUE, K = 20)
save(snspmf.sparse,file = "../snspMF_gwas_analysis/sparsepca.RData")
snspmf.sparse$scree

pca.heatmap.sparse <- pcaPlotMaker(snspmf.sparse$svd$v,joint.mf.gwas,  colnames(joint.mf.gwas), "Singular Vectors, joint PCA (sparse)", ngwas = 55, neqtl = 23)
pca.heatmap.sparse
mix.snspmf.sparse <- mixtureAnalysis(snspmf.sparse$svd$v, colnames(joint.mf.gwas), ngwas = 55, neqtl = 23)
mix.snspmf$plot

pcaBarPlotMaker(snspmf.sparse$svd$v,joint.data.mat =joint.mf.gwas,all.names =colnames(joint.mf.gwas),ngwas = 55, neqtl = 23, "test" )

```
Really not sure of what to make of these.... honestly it seems like there isn't really a signal here at all.Humph

What do the different factors represent?
