---
title: "R Notebook"
output: html_notebook
---

Specific cquestions to answer:
1) Does factor 1 also capture the brain enrichments
2) does GLEANR's sparsity help us better identify genetic architecture?
  Can we identify high selection traits? doo those correspond with those we detect? do they have similar enrichemnts
3) Does GLEANR help us at all in teh blood traits

```{r}
source("/scratch16/abattle4/ashton/snp_networks/gwas_decomp_ldsc/src/plot_functions.R")
source("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization//src/factorEvaluationHelper.R")
source("/scratch16/abattle4/ashton/snp_networks/scratch/factor_interpretation/src/get_pLI.R")

pacman::p_load(magrittr, dplyr, ggplot2, data.table, ggsignif)
f23.tp <- loadGeneEnrichments(23, "top_fe")
f11.tp <- loadGeneEnrichments(11, "top_fe")
f4.tp <- loadGeneEnrichments(4, "top_fe")
```

## Compare LDSC results on factor 1
```{r}

```


## Compare gene set and LDSC enrichments on blood factors:
4,11,23 in GLEANR
3, 33, and 37 in flashR
gleanr enrichments
```{r}
f23.tp <- loadGeneEnrichments(23, "top_fe")
f11.tp <- loadGeneEnrichments(11, "top_fe")
f4.tp <- loadGeneEnrichments(4, "top_fe")
```
flashR enrichments
```{r}
flashr.path = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/results/panUKBB_complete_41K_flashR/top_elements_by_factor/"
flash.f3 <- loadGeneEnrichments(3, "top_fe",path =flashr.path ,unique_only = TRUE)
flash.f33 <- loadGeneEnrichments(33, "top_fe",path =flashr.path ,unique_only = TRUE)
flash.f37 <- loadGeneEnrichments(37, "top_fe",path =flashr.path ,unique_only = TRUE)
```
Okay, what do we see?
It seems like both have cases where they are inflated pretty badly.
```{r}
joined.go_molecular <- left_join(flash.f3$enrichments_unique %>% filter(gene_library=="GO_Molecular_Function_2023"),
f4.tp$enrichments_unique %>% filter(gene_library=="GO_Molecular_Function_2023") %>% select(`Term name`, `P-value`, `Odds ratio`) %>%
  set_colnames(c("Term name", "p_gleanr", "or_gleanr")), by="Term name")

ggplot(joined.go_molecular, aes(x=-log10(`P-value`),y=-log10(p_gleanr))) + geom_point() + geom_abline(intercept = 0,slope=1, col="blue") +
  xlab("flash -log10(p-value)") + ylab("gleanr -log10(p-value)") + theme_bw()
ggplot(joined.go_molecular, aes(x=`Odds ratio`,y=or_gleanr)) + geom_point() + geom_abline(intercept = 0,slope=1, col="blue") +
  xlab("flash OR") + ylab("gleanr OR") + theme_bw()
```
Chromosome location? just of interest
```{r}
joined.go_chr <- left_join(flash.f3$enrichments_unique %>% filter(gene_library=="Chromosome_Location"),
f4.tp$enrichments_unique %>% filter(gene_library=="Chromosome_Location") %>% select(`Term name`, `P-value`, `Odds ratio`) %>%
  set_colnames(c("Term name", "p_gleanr", "or_gleanr")), by="Term name")

ggplot(joined.go_chr, aes(x=-log10(`P-value`),y=-log10(p_gleanr))) + geom_point() + geom_abline(intercept = 0,slope=1, col="blue") +
  xlab("flash -log10(p-value)") + ylab("gleanr -log10(p-value)") + theme_bw()
ggplot(joined.go_chr, aes(x=`Odds ratio`,y=or_gleanr)) + geom_point() + geom_abline(intercept = 0,slope=1, col="blue") +
  xlab("flash OR") + ylab("gleanr OR") + theme_bw()
```

Seems like it happens in cases with just 3 or fewer genes. These are the ones to be careful of.
Look overall at the enrichments of interest, get the correlation:
```{r}
keep.libs <- unique(flash.f3$enrichments_unique$gene_library)[c(2,3,4,5,6,7,8)]
joined.go_all <- left_join(
  flash.f3$enrichments_unique %>% filter(gene_library %in% keep.libs) %>%
    mutate("num_genes_flashr"=str_count(Genes,",")+1),
f4.tp$enrichments_unique %>% filter(gene_library %in% keep.libs) %>%  mutate("num_genes_gleanr"=str_count(Genes,",")+1) %>%
  select(`Term name`, `P-value`, `Odds ratio`,num_genes_gleanr) %>%
  set_colnames(c("Term name", "p_gleanr", "or_gleanr","num_genes_gleanr")), by="Term name")
```
```{r}
joined.go_all <- joined.go_all %>% filter(`Odds ratio` != Inf,or_gleanr != Inf)
ggplot(joined.go_all, aes(x=num_genes_flashr,y=num_genes_gleanr)) + geom_point() + 
  geom_abline(slope = 1,intercept = 0, col="blue") + theme_bw()

ggplot(joined.go_all %>% filter(p_gleanr < 0.001), aes(x=num_genes_flashr,y=num_genes_gleanr)) + 
  geom_point() + geom_abline(slope = 1,intercept = 0, col="blue") + theme_bw()

ggplot(joined.go_all %>% filter(`P-value` < 0.001), aes(x=num_genes_flashr,y=num_genes_gleanr)) + 
  geom_point()+ geom_abline(slope = 1,intercept = 0, col="blue") + theme_bw()


ggplot(joined.go_all, aes(x=`Odds ratio`,y=or_gleanr)) + geom_point() + geom_abline(intercept = 0,slope=1, col="blue") +
  xlab("flash OR") + ylab("gleanr OR") + theme_bw()

ggplot(joined.go_all %>% filter(`P-value` < 0.001 | p_gleanr < 0.001) , aes(x=`Odds ratio`,y=or_gleanr)) + geom_point() + geom_abline(intercept = 0,slope=1, col="blue") +
  xlab("flash OR") + ylab("gleanr OR") + theme_bw()



ggplot(joined.go_all %>% filter(num_genes_flashr >2, num_genes_gleanr > 2) , aes(x=`Odds ratio`,y=or_gleanr)) + geom_point() + geom_abline(intercept = 0,slope=1, col="red", lty="dashed") + geom_smooth(method="lm") + 
  xlab("flash OR") + ylab("gleanr OR") + theme_bw()


joined.go_all.gene_count <- joined.go_all %>% filter(num_genes_flashr >= 3, num_genes_gleanr >= 3)
cor.test(as.numeric(joined.go_all.gene_count$`Odds ratio`), as.numeric(joined.go_all.gene_count$or_gleanr))
cor.test(as.numeric(joined.go_all.gene_count$`P-value`), as.numeric(joined.go_all.gene_count$p_gleanr))
``` 

IS this a similar trend with SVD?
gleanr enrichments
```{r}
f2.gleanr <- loadGeneEnrichments(2, "top_fe")
flashr.path = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/results/panUKBB_complete_41K_flashR/top_elements_by_factor/"
flash.f2 <- loadGeneEnrichments(2, "top_fe",path =flashr.path ,unique_only = TRUE)

svd.path = "/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/results/panUKBB_complete_41K_SVD/top_elements_by_factor/"
svd.f2 <- loadGeneEnrichments(2, "top_fe",path =svd.path ,unique_only = TRUE)
```

```{r}
all_f2 <- left_join(
  flash.f2$enrichments_unique %>% filter(gene_library %in% keep.libs) %>%
    mutate("num_genes_flashr"=str_count(Genes,",")+1) %>%
    select(`Term name`, `P-value`, `Odds ratio`,num_genes_flashr, gene_library) %>%
     set_colnames(c("Term_name", "p_flashr", "or_flashr","num_genes_flashr", "gene_lib")) %>%
    mutate("term_id"=paste0(Term_name, gene_lib)),
  
  svd.f2$enrichments_unique %>% filter(gene_library %in% keep.libs) %>% 
    mutate("num_genes_svd"=str_count(Genes,",")+1) %>%
    mutate("term_id"=paste0(`Term name`, gene_library)) %>%
    select(term_id, `P-value`, `Odds ratio`,num_genes_svd) %>%
     set_colnames(c("term_id", "p_svd", "or_svd","num_genes_svd")),
      by="term_id")

all_f2 <- left_join(all_f2,
                   f2.gleanr$enrichments_unique %>% filter(gene_library %in% keep.libs) %>% 
                  mutate("num_genes_gleanr"=str_count(Genes,",")+1) %>%
                  mutate("term_id"=paste0(`Term name`, gene_library)) %>%
                  select(term_id, `P-value`, `Odds ratio`,num_genes_gleanr) %>%
                   set_colnames(c("term_id", "p_gleanr", "or_gleanr","num_genes_gleanr")),
                    by="term_id")
```

Okay, let's see what we got:
```{r}
cowplot::plot_grid(
ggplot(all_f2, aes(x=num_genes_flashr,y=num_genes_svd)) + geom_point() + theme_bw()+ geom_abline(intercept = 0,slope=1, col="red", lty="dashed")  + geom_smooth(method="lm"),
ggplot(all_f2, aes(x=num_genes_gleanr,y=num_genes_svd)) + geom_point() + theme_bw()+ geom_abline(intercept = 0,slope=1, col="red", lty="dashed") + geom_smooth(method="lm"),
ggplot(all_f2, aes(x=num_genes_gleanr,y=num_genes_flashr)) + geom_point() + theme_bw()+ geom_abline(intercept = 0,slope=1, col="red", lty="dashed") + geom_smooth(method="lm"), nrow=1)
```
Okay- both flashR and SVd get more genes in their sets. This is unsurprising, given that the factors are much more dense.
Does this result in improved power?
```{r}
cowplot::plot_grid(
ggplot(all_f2 %>% filter(num_genes_flashr > 3,num_genes_svd > 3 ), aes(x=-log10(p_flashr),y=-log10(p_svd))) + geom_point() + theme_bw()+ geom_abline(intercept = 0,slope=1, col="red", lty="dashed")  + geom_smooth(method="lm"),
ggplot(all_f2 %>% filter(num_genes_gleanr > 3,num_genes_svd > 3 ), aes(x=-log10(p_gleanr),y=-log10(p_svd))) + geom_point() + theme_bw()+ geom_abline(intercept = 0,slope=1, col="red", lty="dashed")  + geom_smooth(method="lm"),
ggplot(all_f2 %>% filter(num_genes_flashr > 3,num_genes_gleanr > 3 ), aes(x=-log10(p_gleanr),y=-log10(p_flashr))) + geom_point() + theme_bw()+ geom_abline(intercept = 0,slope=1, col="red", lty="dashed")  + geom_smooth(method="lm"),nrow=1)
```


## Compare teh genetic architecture:
Make similar plots- do they stratify as nicely?
GLEANR factors 18, 32,37,43
flash factors 16, 29,44,18



```{r}
selection.and.polygen.flashr <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/results/panUKBB_complete_41K_flashR//selective_pressure/s_scores.tsv")

selection.and.polygen.svd <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/results/panUKBB_complete_41K_SVD//selective_pressure/s_scores.tsv")

selection.and.polygen.svd_scaled <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/results/panUKBB_complete_41K_SVD_scaled/selective_pressure/s_scores.tsv")

selection.and.polygen.gleanr <- fread("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/results/panUKBB_complete_41K_final//selective_pressure/s_scores.tsv")

head(selection.and.polygen.flashr)
```

```{r}
flash <- ggplot(selection.and.polygen.flashr, aes(x=Me_scaled,y=S_hat)) + geom_point() + theme_bw() + ggtitle("flash")

svd <- ggplot(selection.and.polygen.svd, aes(x=Me_scaled,y=S_hat)) + geom_point() + theme_bw() + ggtitle("SVD")
svd_scaled <- ggplot(selection.and.polygen.svd_scaled, aes(x=Me_scaled,y=S_hat)) + geom_point() + theme_bw() + ggtitle("SVD (scaled)")

gleanr <- ggplot(selection.and.polygen.gleanr, aes(x=Me_scaled,y=S_hat)) + geom_point()+ theme_bw() + ggtitle("gleanr")

cowplot::plot_grid(gleanr, flash, svd,svd_scaled)
```

