---
title: "Supplemental figures for special cases ."
output: html_notebook
---
Here, creating the simulation based figure
Note that previous sims were incorrect in that they didn't quite use the sd specifications in the right way. We should be able to get this to work
## Update the paths below to reflect the new locations, things have been mooved about!"
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/gwas_sim_style/simulation_processing_tools.R")
source("/scratch16/abattle4/ashton/snp_networks/gwas_decomp_ldsc/src/plot_functions.R")
source("/scratch16/abattle4/ashton/snp_networks/gwas_decomp_ldsc/src/factorization_methods.R")
pacman::p_load(data.table, magrittr, dplyr,stringr, ggplot2, RColorBrewer)

```

### Special cases 1-3: No dense factors, many dense factors, and some "denser" factors
```{r}
no.covar <- mergeTabularResults(query="",
                                base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_no_overlap/",
                                quiet=TRUE)%>% mutate("Covar"=  "none")


oneb.covar <- mergeTabularResults(query="",
                                  base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_1b_overlap/",
                                  quiet=TRUE) %>% mutate("Covar"=  "1b")

twob.covar <- mergeTabularResults(query="",
                                  base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_2b_overlap//",
                                  quiet=TRUE) %>% mutate("Covar"=  "2b")
```

### Special case 3: some *denser* factors

```{r}
no.covar.nonsparse <- mergeTabularResults(query="",
                                base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_mid-density_no_overlap//",
                                quiet=TRUE)%>% mutate("Covar"=  "none")


oneb.covar.nonsparse <- mergeTabularResults(query="",
                                  base.dir ="/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_mid-density_1b_overlap//",
                                  quiet=TRUE) %>% mutate("Covar"=  "1b")

twob.covar.nonsparse <- mergeTabularResults(query="",
                                  base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_mid-density_2b_overlap//",
                                  quiet=TRUE) %>% mutate("Covar"=  "2b")
```

Make sure to get the N set in the right order
```{r}
n.levels = c("5000","10000","50000","1e+05","2e+05","mixed")
joined.special.cases <- rbind(no.covar, oneb.covar, twob.covar,no.covar.nonsparse,oneb.covar.nonsparse,twob.covar.nonsparse) %>% 
  select(-R2_L, -R2_F,-joint_pr_auc_v,-joint_pr_auc_u,-joint_pr_kappa_v,-joint_pr_kappa_u,-joint_pr_pearson_v,-joint_pr_pearson_u) %>%
  rowwise() %>%
  mutate("V_mat"=stringr::str_match(fname,pattern="^(V\\d+)_U")[2],
         "U_mat"=stringr::str_match(fname,pattern="V\\d+_(U\\d+)_MAF")[2]) %>% 
  mutate("test_type"=case_when(V_mat == "V105" ~"no_dense",
                               V_mat == "V106" ~ "no_dense",
                               V_mat == "V107" ~ "3_dense",
                               V_mat == "V109" ~ "5_dense",
                               V_mat == "V110" ~ "low_sparsity",
                               V_mat == "V111" ~ "low_sparsity")) %>% ungroup()
head(joined.special.cases)
```

## Aesthetic ordering and coloration elements
Build the color pallete
```{r}
"new.method"=c("SVD (avg)", "SVD (elbow)", "SVD-adj (avg)","SVD-adj (elbow)","flash", "GLEANR-U","GLEANR")

full.list=c("SVD", "SVD (avg)", "SVD (elbow)","SVD-adj", "SVD-adj (avg)","SVD-adj (elbow)",
            "flash","flash (Z)","flash (Kronecker)","FactorGo","GLEANR-U","GLEANR")
flash.cols <- brewer.pal(5,"Greens")[3:5]
pca.unadj.cols <- brewer.pal(3,"Blues")
pca.adj.cols <- brewer.pal(8,"BuPu")[c(5,6,8)]
pca.adj.cols <- brewer.pal(3,"Purples")
GLEANR.cols <- brewer.pal(2,"Oranges")[2:3]
factorgo.cols <- brewer.pal(5,"PuRd")[4]

all.colors <- c(pca.unadj.cols,pca.adj.cols, flash.cols, factorgo.cols,GLEANR.cols )
color.map.pub <- data.frame("method_label"=full.list, "color" = all.colors)

myColors <- color.map.pub$color
names(myColors) <- levels(cons.matrix$method_label)
colScale <- scale_colour_manual(name = "Method",values = myColors)
fillScale <- scale_fill_manual(name = "Method",values = myColors)
```

```{r}
name.map = data.frame("method_label"=c("SVD", "flash (Z)", "flash", "flash (Kronecker)","SVD (avg)", "GLEANR", "GLEANR-U", "SVD-adj", "SVD (beta)", "FactorGo"), "method" = unique(cons.matrix$method))
level_order=name.map$method_label
```
Apply the colors:
```{r}
cons.matrix %<>% left_join(.,name.map, by="method") %>% 
  mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% 
  arrange(method_label) %>% filter(method != "SVD_beta",method_label != "flash (Z)")

cons.factor %<>% left_join(.,name.map, by="method") %>% 
  mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% 
  arrange(method_label) %>% filter(method != "SVD_beta",method_label != "flash (Z)")
  
cons.entry %<>% left_join(.,name.map, by="method") %>% 
  mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% 
  arrange(method_label) %>% filter(method != "SVD_beta",method_label != "flash (Z)")
```
Clean up labels
```{r}
pop_names <- as_labeller(
     c(`5000` = "N=5k", `10000` = "N=10k",`50000` = "N=50k", 
       `1e+05` = "N=100k",`2e+05` = "N=200k", "mixed"="N=[10K,200K]",
     `1b`="1-block", `2b`="2-block", `none`="No overlap"))

# Supplemental figure
drop.methods <- c("flash (Z)", "SVD (elbow)","SVD (elbow)","SVD-adj (elbow)","SVD-adj (avg)", "SVD (avg)")
```

### Make the factor plots
#### Per factor variance
V
```{r}
v.sim.factor <- ggplot(cons.factor %>% filter((!method_label %in% drop.methods), matrix == "V"), 
aes(x = method_label, y = factor_vars, fill = method_label)) + 
  geom_boxplot() + facet_grid(Covar~N,labeller = pop_names) +  theme_bw() + xlab("Factorization method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + ylab("Variance across factors") + 
  fillScale +  theme(strip.background =element_rect(fill="white"),legend.position = "top") + 
  ggtitle("Variance across simulations runs with respect to V")
v.sim.factor

v.leg <- ggpubr::get_legend(v.sim.factor)
```
U:
```{r}
u.sim.factor <- ggplot(cons.factor %>% filter((!method_label %in% drop.methods), matrix == "U"), 
                aes(x = method_label, y = factor_vars, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(14) +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab(bquote(R^2 ~ "with simulated U")) +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left") +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + 
  theme(legend.text=element_text(size=12)) + 
  theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt")) + 
  ggtitle("Covariance adjustment improves reconstruction of simulated genetic factors") + 
  theme(plot.title = element_text(size=10,color="dimgrey"))  
```

#### Per entry variance
V
```{r}
v.sim.entry <- ggplot(cons.entry %>% filter((!method_label %in% drop.methods), matrix == "V"), 
aes(x = method_label, y = entry_var, fill = method_label)) + 
  geom_boxplot() + facet_grid(Covar~N,labeller = pop_names) +  theme_bw() + xlab("Factorization method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + ylab("Variance across entries") + 
  fillScale +  theme(strip.background =element_rect(fill="white"),legend.position = "top") + 
  ggtitle("Variance across simulations runs with respect to V")
v.sim.entry
```
U:
```{r}
u.sim.entry <- ggplot(cons.entry %>% filter((!method_label %in% drop.methods), matrix == "U"), 
                aes(x = method_label, y = entry_var, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(14) +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab("Variance per entry") +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left") +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + 
  theme(legend.text=element_text(size=12)) + 
  theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt")) + 
  ggtitle("Covariance adjustment improves reconstruction of simulated genetic factors") + 
  theme(plot.title = element_text(size=10,color="dimgrey"))  

u.sim.entry
```

#### Per matrix
V
```{r}
v.sim.matrix <- ggplot(cons.matrix %>% filter((!method_label %in% drop.methods), matrix == "V"), 
aes(x = method_label, y = matrix_vars, fill = method_label)) + 
  geom_boxplot() + facet_grid(Covar~N,labeller = pop_names) +  theme_bw() + xlab("Factorization method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + ylab("Variance across matrices") + 
  fillScale +  theme(strip.background =element_rect(fill="white"),legend.position = "top") + 
  ggtitle("Variance across simulations runs with respect to V")
v.sim.matrix
```
U:
```{r}
u.sim.matrix <- ggplot(cons.matrix %>% filter((!method_label %in% drop.methods), matrix == "U"), 
                aes(x = method_label, y = matrix_vars, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(14) +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab("Variance across matrices") +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left") +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + 
  theme(legend.text=element_text(size=12)) + 
  theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt")) + 
  ggtitle("Covariance adjustment improves reconstruction of simulated genetic factors") + 
  theme(plot.title = element_text(size=10,color="dimgrey"))  

u.sim.matrix
```
