---
title: "Supplemental figures for special cases ."
output: html_notebook
---
Here, creating the simulation based figure
Note that previous sims were incorrect in that they didn't quite use the sd specifications in the right way. We should be able to get this to work
## Update the paths below to reflect the new locations, things have been mooved about!"
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/gwas_sim_style/simulation_processing_tools.R")
source("/scratch16/abattle4/ashton/snp_networks/gwas_decomp_ldsc/src/plot_functions.R")
source("/scratch16/abattle4/ashton/snp_networks/gwas_decomp_ldsc/src/factorization_methods.R")
pacman::p_load(data.table, magrittr, dplyr,stringr, ggplot2, RColorBrewer)

```

### Special cases 1,2: No dense factors, some "denser" factors
```{r}
no.covar <- mergeTabularResults(query="",
                                base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_no_overlap/",
                                quiet=TRUE)%>% mutate("Covar"=  "none")


oneb.covar <- mergeTabularResults(query="",
                                  base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_1b_overlap/",
                                  quiet=TRUE) %>% mutate("Covar"=  "1b")

twob.covar <- mergeTabularResults(query="",
                                  base.dir = "/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_2b_overlap//",
                                  quiet=TRUE) %>% mutate("Covar"=  "2b")
```
#### Special case 3: Multiple dense factors
```{r}
bd="/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/"
no.covar.dense <- mergeTabularResults(query="",
                                base.dir = paste0(bd, "dense_no_overlap/"),
                                quiet=TRUE)%>% mutate("Covar"=  "none")


oneb.covar.dense <- mergeTabularResults(query="",
                                  base.dir = paste0(bd, "dense_1b_overlap/"),
                                  quiet=TRUE) %>% mutate("Covar"=  "1b")

twob.covar.dense <- mergeTabularResults(query="",
                                  base.dir = paste0(bd, "dense_2b_overlap/"),
                                  quiet=TRUE) %>% mutate("Covar"=  "2b")
```

Make sure to get the N set in the right order
```{r}
n.levels = c("5000","10000","50000","1e+05","2e+05","mixed")
joined.special.cases <- rbind(no.covar, oneb.covar, twob.covar,no.covar.dense,oneb.covar.dense,twob.covar.dense) %>% 
  select(-R2_L, -R2_F,-joint_pr_auc_v,-joint_pr_auc_u,-joint_pr_kappa_v,-joint_pr_kappa_u,-joint_pr_pearson_v,-joint_pr_pearson_u) %>%
  rowwise() %>%
  mutate("V_mat"=stringr::str_match(fname,pattern="^(V\\d+)_U")[2],
         "U_mat"=stringr::str_match(fname,pattern="V\\d+_(U\\d+)_MAF")[2]) %>% 
  mutate("test_type"=case_when(V_mat == "V105" ~"no_dense",
                               V_mat == "V106" ~ "no_dense",
                               V_mat == "V107" ~ "sparse_bonus_run",
                               V_mat == "V109" ~ "sparse_bonus_run",
                               V_mat %in% c("V113", "V115")  ~ "all_dense",
                               V_mat %in% c("V114", "V116")  ~ "3_dense",
                               V_mat %in% c("V110", "V111")  ~ "low_sparsity")) %>% 
           ungroup() %>% filter(V_mat != "V107",V_mat != "V109" )
#these bonus runs were generated inadvertently, don't need to be included.
head(joined.special.cases)
```

A few sanity checks:
```{r}
unique(joined.special.cases$method)
#10. methods x 10 sims each
all((joined.special.cases %>% group_by(fname,) %>% summarize("count"=n()))$count == 100)
```

## Aesthetic ordering and coloration elements
Make sure to get the N set in the right order
```{r}
n.levels = c("5000","10000","50000","1e+05","2e+05","mixed")
joined.special.cases %<>%  mutate("N"=factor(N,levels=n.levels))
```
Make sure the names are updated to be clean:
```{r}
name.map = data.frame("method_label"=c("SVD", "flash (Z)", "flash", "flash (Kronecker)","SVD (avg)", "GLEANR", "GLEANR-U", "SVD-adj", "SVD (beta)", "FactorGo"), "method" = unique(joined.special.cases$method))
level_order=name.map$method_label
```

Build the color pallete
```{r}
color.map.pub <-buildPubColors()
```
Apply the colors:
```{r}
joined.special.cases %<>% left_join(.,name.map, by="method") %>% 
  mutate("method_label"=factor(method_label, levels=color.map.pub$method_label)) %>% 
  arrange(method_label) %>% filter(method != "SVD_beta",method_label != "flash (Z)")
```
Clean up labels
```{r}
pop_names <- as_labeller(
     c(`5000` = "N=5k", `10000` = "N=10k",`50000` = "N=50k", 
       `1e+05` = "N=100k",`2e+05` = "N=200k", "mixed"="N=[10K,200K]",
     `1b`="1-block", `2b`="2-block", `none`="No overlap"))

# Supplemental figure
drop.methods <- c("flash (Z)", "SVD (elbow)","SVD (elbow)","SVD-adj (elbow)","SVD-adj (avg)", "SVD (avg)")

#build the color list:

col.list <- buildColorList(color.map.pub,joined.special.cases$method_label)
fillScale <- col.list$fillScale
colScale <- col.list$colScale
```

### Make the factor plots
```{r}
core.dat.plt <- joined.special.cases %>% select(method, yuan_U, yuan_V, BIC,MAF,N,fname, Covar,V_mat, U_mat, test_type, method_label)
```
### Debugging a weird issue:
Previously we observed no deviation in performance in GLEANR and GLEANR-U, due to an error in the yaml settings file.
This has been repaired.
```{r}
nocovar <- core.dat.plt %>%filter(method %in% c("GLEANER_glmnet_noCovar")) %>% select(method, yuan_U,fname,BIC) %>% arrange(fname, BIC)
covar <- core.dat.plt %>%filter(method %in% c("GLEANER_glmnet")) %>% select(method, yuan_U,fname,BIC) %>% arrange(fname, BIC)
stopifnot(all(nocovar$fname == covar$fname))
plot(nocovar$yuan_U, covar$yuan_U); abline(a=0,b=1,col="blue")
```

#### Case 1: no dense factors
Number of sims
```{r}
length(unique((core.dat.plt %>% filter(test_type == "no_dense"))$fname)) * 10
```

V
```{r}
v.sim.no_dense <- ggplot(core.dat.plt %>% filter(test_type == "no_dense")  %>% filter(!(method_label %in% drop.methods)), 
aes(x = method_label, y = yuan_V, fill = method_label)) + 
  geom_boxplot() + facet_grid(Covar~N,labeller = pop_names) +  theme_bw() + xlab("Factorization method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + ylab(bquote(R^2~"wrt V")) + 
  fillScale +  theme(strip.background =element_rect(fill="white"),legend.position = "top") + 
  ggtitle(bquote(R^2 ~ "with respect to V (no dense factors)"))

v.sim.no_dense
```
Numerically, what do we see?
```{r}
core.dat.plt %>% filter(test_type == "no_dense") %>% group_by(method_label, N, Covar) %>% summarize("median_u"=median(yuan_U),"median_v"=median(yuan_V)) %>%
  ungroup() %>% group_by(N,Covar) %>% slice_max(median_v) %>% arrange(N)
```

U:
```{r}
u.sim.no_dense <- ggplot(core.dat.plt %>% filter(test_type == "no_dense")  %>% filter(!(method_label %in% drop.methods)), 
aes(x = method_label, y = yuan_U, fill = method_label)) + 
  geom_boxplot() + facet_grid(Covar~N,labeller = pop_names) +  theme_bw() + xlab("Factorization method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + ylab(bquote(R^2~"wrt U")) + 
  fillScale +  theme(strip.background =element_rect(fill="white"),legend.position = "top") + 
  ggtitle(bquote(R^2 ~ "with respect to U (no dense factors in V)"))

u.sim.no_dense
```
Numerically, we see that in a
```{r}
core.dat.plt %>% filter(test_type == "no_dense") %>% group_by(fname,method) %>% summarize("median_u"=median(yuan_U),"median_v"=median(yuan_V)) %>%
  ungroup() %>% group_by(fname) %>% slice_max(median_u) 


core.dat.plt %>% filter(test_type == "no_dense") %>% group_by(method_label, N, Covar) %>% summarize("median_u"=median(yuan_U),"median_v"=median(yuan_V)) %>%
  ungroup() %>% group_by(N,Covar) %>% slice_max(median_u) %>% arrange(N)
```

#### 2) multiple dense factors
How many are there??
```{r}
length(unique((core.dat.plt %>% filter(test_type == "3_dense"))$fname))
```

V
```{r}
v.sim.3_dense <- ggplot(core.dat.plt %>% filter(test_type == "3_dense") %>% filter(!(method_label %in% drop.methods)), 
aes(x = method_label, y = yuan_V, fill = method_label)) + 
  geom_boxplot() + facet_grid(Covar~N,labeller = pop_names) +  theme_bw() + xlab("Factorization method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + ylab(bquote(R^2~"wrt V")) + 
  fillScale +  theme(strip.background =element_rect(fill="white"),legend.position = "top") + 
  ggtitle(bquote(R^2 ~ "with respect to V (3 dense factors)"))

v.sim.3_dense
```
Dig in to this, our good performance is unexpected:
Look at iter 10
```{r}
tabular.performance <- fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_no_overlap/V107_U103_MAF-mix_eur_N-50000_RHO-none_No-none/factorization_results/summary.tabular_performance.tsv") %>% arrange(-yuan_V)
tabular.performance

gleanr.v <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_no_overlap/V107_U103_MAF-mix_eur_N-50000_RHO-none_No-none/factorization_results/sim10.GLEANER_glmnet.factors.txt"))
gleanr.u <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_no_overlap/V107_U103_MAF-mix_eur_N-50000_RHO-none_No-none/factorization_results/sim10.GLEANER_glmnet.loadings.txt"))

#Flash_SE is the better performing one it looks like. Verify this behaves as intended
flashr.v <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_no_overlap/V107_U103_MAF-mix_eur_N-50000_RHO-none_No-none/factorization_results/sim10.FLASH_SE.factors.txt"))
flashr.u <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/simulation_outputs/final_sims_june_2024/special_no_overlap/V107_U103_MAF-mix_eur_N-50000_RHO-none_No-none/factorization_results/sim10.FLASH_SE.loadings.txt"))

ref.v <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/u_and_v/V107_M10_K5_3dense.csv"))
ref.v.full <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/u_and_v/V109_M10_K5_5dense.csv"))
ref.u <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/u_and_v/U103_N100_K5.csv"))
```

Okay, step back. Let's make sure all the new sims match our expectations.



```{r}
source("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/src/matrix_similarity_scoring/evaluateSimR2.R")
source("/scratch16/abattle4/ashton/snp_networks/custom_l1_factorization/src/matrix_similarity_scoring/matrix_comparison_utils.R")

ref_v_gleanr <- evaluate_error(unitNorms(ref.u), unitNorms(ref.v), unitNorms(gleanr.u), unitNorms(gleanr.v)) #Why is this so much longer than the std version?
ref_v_flashr <- evaluate_error(unitNorms(ref.u), unitNorms(ref.v), unitNorms(flashr.u), unitNorms(flashr.v)) #Why is this so much longer than the std version?

cowplot::plot_grid(
  plotFactors(as.matrix(unitNorms(ref.v)), trait_names = paste0("T", 1:10), title="Ref V",cluster=FALSE),
plotFactors(as.matrix(ref_v_gleanr$reordered_V), trait_names = paste0("T", 1:10), title="GLEANR V",cluster = FALSE),
plotFactors(as.matrix(ref_v_flashr$reordered_V), trait_names = paste0("T", 1:10), title="flashR V",cluster = FALSE),
nrow=1)

```

U:
```{r}
u.sim.entry <- ggplot(cons.entry %>% filter(!(method_label %in% drop.methods), matrix == "U"), 
                aes(x = method_label, y = entry_var, fill = method_label)) +
  geom_boxplot() + 
  facet_grid(Covar~N,labeller = pop_names) +  theme_bw(14) +xlab("Factorization method")  + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + 
  ylab("Variance per entry") +
  fillScale + theme(strip.background =element_rect(fill="white"), legend.position = "left") +
  guides(fill = guide_legend(ncol = 2, 
                             override.aes = list(shape = 22, 
                                                 size = 8, 
                                                 linetype = 0, 
                                                 color = NA))) + 
  theme(legend.text=element_text(size=12)) + 
  theme(legend.position="none",margin(t = 0, r = -1, b = 0, l = 0, unit = "pt")) + 
  ggtitle("Covariance adjustment improves reconstruction of simulated genetic factors") + 
  theme(plot.title = element_text(size=10,color="dimgrey"))  

u.sim.entry
```

#Pause on the above, need to revisit and fix.

### Factors with middling density.
Visualize them for sanity's sake (sims 110 and 111):
```{r}
ref.v.mid1 <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/u_and_v/V110_M10_K5_mid-sparse.csv"))
ref.v.mid2 <- as.matrix(fread("/scratch16/abattle4/ashton/snp_networks/scratch/cohort_overlap_exploration/simulating_factors/custom_easy/u_and_v/V111_M10_K5_mid-sparse.csv"))
```
Plot

```{r}
plotFactors(ref.v.mid1, trait_names = paste0("T", 1:10), title="Init mat",cluster = FALSE)
ref.v.mid1
plotFactors(ref.v.mid2, trait_names = paste0("T", 1:10), title="Init mat",cluster = FALSE)
ref.v.mid2
```
Great, that's fine.
```{r}
v.sim.mid_dense <- ggplot(core.dat.plt %>% filter(test_type == "low_sparsity") %>% filter(!(method_label %in% drop.methods)), 
aes(x = method_label, y = yuan_V, fill = method_label)) + 
  geom_boxplot() + facet_grid(Covar~N,labeller = pop_names) +  theme_bw() + xlab("Factorization method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + ylab(bquote(R^2~"wrt V")) + 
  fillScale +  theme(strip.background =element_rect(fill="white"),legend.position = "top") + 
  ggtitle("R2 with respective to V with low sparsity factors")

v.sim.mid_dense
```
U:
```{r}
u.sim.mid_dense <- ggplot(core.dat.plt %>% filter(test_type == "low_sparsity") %>% filter(!(method_label %in% drop.methods)), 
aes(x = method_label, y = yuan_U, fill = method_label)) + 
  geom_boxplot() + facet_grid(Covar~N,labeller = pop_names) +  theme_bw() + xlab("Factorization method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) + ylab(bquote(R^2~"wrt U")) + 
  fillScale +  theme(strip.background =element_rect(fill="white"),legend.position = "top") + 
  ggtitle("R2 with respective to U with low sparsity factors")

u.sim.mid_dense
```
